### 2015-01-06 prelim/oral_prelim.tex
moving files around for oral prelim

\begin{frame}{Multivariate analysis}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item computationally challenging to work with
      \item joint distribution only available in low dimension
    \end{itemize}
    \item Bayesian hierarchical model
    \item Pairwise likelihood approach (Huser and Davison, 2014)
  \end{itemize}
\end{frame}

@@

\begin{frame}{Thresholding data}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item We threshold the observed data at a high threshold $T$.
    \item Thresholded data:
    \begin{align*}
      Y_t^*(\bs) = \left\{ \begin{array}{ll}
          Y_t(\bs) \quad & Y_t(\bs) > T\\
          T \quad & Y_t(\bs) \le T
      \end{array}\right.
    \end{align*}
    \item Allows tails of the distribution to speak for themselves.
  \end{itemize}
\end{frame}

\begin{frame}{$\chi$ coefficient}
  \begin{itemize} \setlength{\itemsep}{0.5em}
   \item The $\chi$ coefficient is a measure of extremal dependence
   \item Specifically, we focus on $\chi(\bh)$ for the upper tail given by
    \begin{align*}
      \chi(\bh) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bs + \bh) > c)
    \end{align*}
    \item If $ \chi(\bh) = 0$, then observations are asymptotically independent at distance $\bh$.
    \item We expect $\lim_{\bh \rightarrow \infty}\chi(\bh)=0$.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Example daily partition}
    \centering
    \begin{figure}
    \includegraphics[width=0.54\linewidth]{./plots/example-partition-1.pdf}
    \includegraphics[width=0.54\linewidth]{./plots/example-partition-2.pdf}
    \caption{Two sample partitions (number is at partition center)}
    \end{figure}
\end{frame}

\begin{frame}{Simulated $\widehat{\chi}(\bh)$ plots}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/chi-plots.pdf}\\[-0.25in]
  \includegraphics[width=0.2\linewidth]{./plots/chi-legend.pdf}
  \caption{$\chi$ plots from simulated datasets}
  \end{figure}
\end{frame}

\begin{frame}{Sample simulated datasets}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/gauss-vs-t3.pdf}
  \caption{Gaussian and $t$ with 3 partitions}
  \end{figure}
\end{frame}

\begin{frame}{Sample simulated datasets}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/gauss-vs-skew-t3.pdf}
  \caption{Gaussian and skew-$t$ with 3 partitions}
  \end{figure}
\end{frame}

@@

\begin{frame}{Data analysis}
\begin{columns}[c]
\column{.45 \linewidth}
	\begin{itemize} \setlength{\itemsep}{0.5em}
	\item Data analysis uses
	\begin{itemize}
		\item max 8-hour ozone measurements
		\item 85 sites
		\item 92 days
	\end{itemize}
	\end{itemize}

	\column{.55\linewidth}
	\begin{figure}
    \includegraphics[width=1\linewidth]{./plots/ozone_station.pdf}
    \caption{Ozone monitoring station locations}
    \end{figure}
\end{columns}
\end{frame}

\begin{frame}{Data analysis}
  \centering
  \begin{figure}
    \includegraphics[width=1\linewidth]{./plots/ozone-day.pdf}
    \caption{Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for days 5 and 34}
   \end{figure}

@@

\begin{frame}{Exploratory data analysis}
	\centering
  \begin{figure}
    \includegraphics[width=1\linewidth]{./plots/chi-plot-ozone-res.pdf}
    \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
  \end{figure}
\end{frame}

@@

\begin{frame}{Quantile score for cross-validation}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item The quantile score for the $\tau$th quantile is
    \begin{align*}
      2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
    \end{align*}
    where:
    \begin{itemize}
      \item $y$ is a test set value
      \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Five-fold cross-validation results}
  \begin{table}[htbp]
    \small
    \centering
    \begin{tabular}{l|c|r|rrrrr}
          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{$\tau$}\\
           \hline
  Marginal & $K$ & $T$  & 0.950 & 0.980 & 0.990 & 0.995 & 0.999\\
  \hline
Gaussian 	& 1 & 0 	& 39.820 & 17.539 & 9.167 & 4.720 & 1.057\\
$t$		& 1 & 0 	& {\bf 31.008} & {\bf 13.898} & 7.229 & {\bf 3.405} & 0.879\\
$t$		& 3 & 0	& 31.213 & 13.920 & {\bf 7.218} & 3.498 & 0.918\\
$t$		& 1 & 0.9	& 32.221 & 14.519 & 7.549 & 3.604 & 0.896\\
$t$		& 3 & 0.9 	& 38.842 & 16.781 & 8.434 & 4.180 & 1.020\\
skew-$t$	& 1 & 0	& 31.845 & 14.542 & 7.533 & 3.645 & {\bf 0.844}\\
skew-$t$	& 1 & 0.9	& 32.132 & 14.296 & 7.484 & 3.497 & 0.890\\
skew-$t$	& 3 & 0	& 33.653 & 15.453 & 8.119 & 4.338 & 1.188\\
skew-$t$	& 3 & 0.9 	& 32.157 & 14.727 & 7.794 & 3.825 & 0.917\\
\hline
    \end{tabular}
    \caption{Brier score for predicting exceedance of $c = \hat{q}(\tau)$ from five-fold cross-validation ($\times 1000$)}
  \end{table} \vspace{-1.5em}
  \begin{itemize}
  	\item Quantile score results are similar
  \end{itemize}
\end{frame}

\begin{frame}{Predicted 95th quantile}
\centering
\begin{figure}
    \includegraphics[width=.5\linewidth]{./plots/quantile-95-gau.pdf}
    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
    \caption{Predicted 95th quantile using Gaussian and $t$}
\end{figure}
\end{frame}

\begin{frame}{Predicted 95th quantile}
\centering
\begin{figure}
    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t19.pdf}
    \caption{Predicted 95th quantile using $t$ and $t$ thresholded at $T=0.9$}
\end{figure}
\end{frame}

\begin{frame}{Predicted 99th quantile}
\centering
\begin{figure}
    \includegraphics[width=.5\linewidth]{./plots/quantile-99-gau.pdf}
    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
    \caption{Predicted 99th quantile using Gaussian and $t$}
\end{figure}
\end{frame}

\begin{frame}{Predicted 99th quantile}
\centering
\begin{figure}
    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t19.pdf}
    \caption{Predicted 99th quantile using $t$ and $t$ thresholded at $T=0.9$}
\end{figure}
\end{frame}

\begin{frame}{Probability of exceedance}
\centering
\begin{figure}
    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-gau.pdf}
    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
    \caption{Probability of exceeding the 75 ppb ozone standard using Gaussian and $t$}
\end{figure}
\end{frame}

\begin{frame}{Probability of exceedance}
\centering
\begin{figure}
    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t19.pdf}
    \caption{Probability of exceeding the 75 ppb ozone standard using $t$ and $t$ thresholded at $T=0.9$}
\end{figure}
\end{frame}

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 6 different data settings:
    \begin{itemize}
    	\item Gaussian vs $t$ vs skew-$t$ marginal distribution
        \item $K=1$ partition vs $K=5$ partitions
    \end{itemize}
    \item Preliminary results are inconclusive
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Multivariate analysis}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item computationally challenging to work with
      \item joint distribution only available in low dimension
    \end{itemize}
    \item Bayesian hierarchical model (Reich and Shaby, 2012)
    \item Pairwise likelihood approach (Huser and Davison, 2014)
  \end{itemize}
\end{frame}

@@

\begin{frame}{Censoring data}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item We censor the observed data at a high threshold $T$.
    \item Censored data:
    \begin{align*}
      \tilde{Y}_t(\bs) = \left\{ \begin{array}{ll}
          Y_t(\bs) \quad & \delta(\bs) = 1\\
          T \quad & \delta(\bs) = 0
      \end{array}\right.
    \end{align*}
    where $\delta(\bs) = I[Y(\bs) > T]$
    \item Allows tails of the distribution to speak for themselves.
  \end{itemize}
\end{frame}

\begin{frame}{$\chi$ coefficient}
  \begin{itemize} \setlength{\itemsep}{0.5em}
   \item The $\chi$ coefficient is a measure of extremal dependence
   \item Specifically, we focus on $\chi(\bh)$ for the upper tail given by
    \begin{align*}
      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
    \end{align*}
    where $h = ||\bs - \bt||$
    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $\bh$.
    \item We expect $\lim_{\bh \rightarrow \infty}\chi(\bh)=0$.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Example daily partition}
    \centering
    \begin{figure}
    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-1.pdf}
    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-2.pdf}
    \caption{Two sample partitions (number is at partition center)}
    \end{figure}
\end{frame}

\begin{frame}{Simulated $\widehat{\chi}(\bh)$ plots}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/pot/chi-h.pdf}\\[-0.25in]
  \caption{$\chi$ plot for different data settings}
  \end{figure}
\end{frame}

\begin{frame}{Sample simulated datasets}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-t3.pdf}
  \caption{Gaussian and $t$ with 3 partitions}
  \end{figure}
\end{frame}

\begin{frame}{Sample simulated datasets}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-skew-t3.pdf}
  \caption{Gaussian and skew-$t$ with 3 partitions}
  \end{figure}
\end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 6 different data settings:
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
        \item $K=1$ partition vs $K=5$ partitions
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Brier score results}
\centering

@@

\begin{frame}{Data analysis}
\begin{columns}[c]
\column{.45 \linewidth}
	\begin{itemize} \setlength{\itemsep}{0.5em}
	\item Ozone measurements
	\begin{itemize}
		\item max 8-hour ozone measurements
		\item data from 1089 sites
		\item July 2005
	\end{itemize}
  \item We take a stratified sample of $n = 800$ sites:
  \begin{itemize}
    \item 271 from northeast
    \item 96 from northwest
    \item 269 from southeast
    \item 164 from southwest
  \end{itemize}
	\end{itemize}

	\column{.55\linewidth}
	\begin{figure}
    \includegraphics[width=1\linewidth]{./plots/pot/ozone_stations.pdf}
    \caption{Ozone monitoring station locations}
    \end{figure}
\end{columns}
\end{frame}

\begin{frame}{Data analysis}
  \centering
  \begin{figure}
    \includegraphics[width=1\linewidth]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}

\begin{frame}{Exploratory data analysis}
	\centering
  \begin{figure}
    \includegraphics[width=1\linewidth]{./plots/pot/chi-plot-ozone-res.pdf}
    \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
  \end{figure}
\end{frame}

@@

% \begin{frame}{Quantile score for cross-validation}
%   \begin{itemize} \setlength{\itemsep}{0.5em}
%     \item The quantile score for the $\tau$th quantile is
%     \begin{align*}
%       2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
%     \end{align*}
%     where:
%     \begin{itemize}
%       \item $y$ is a test set value
%       \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
%     \end{itemize}
%   \end{itemize}
% \end{frame}

@@

\begin{frame}{Five-fold cross-validation results}
  \begin{table}[htbp]
    \small
    \centering
    \begin{tabular}{l|c|r|rrrrr}
          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{$\tau$}\\
           \hline
  Marginal & $K$ & $T$  & 0.950 & 0.980 & 0.990 & 0.995 & 0.999\\
  \hline
Gaussian 	& 1 & 0 	& 39.820 & 17.539 & 9.167 & 4.720 & 1.057\\
$t$		& 1 & 0 	& {\bf 31.008} & {\bf 13.898} & 7.229 & {\bf 3.405} & 0.879\\
$t$		& 5 & 0	& 31.213 & 13.920 & {\bf 7.218} & 3.498 & 0.918\\
$t$		& 1 & 0.9	& 32.221 & 14.519 & 7.549 & 3.604 & 0.896\\
$t$		& 5 & 0.9 	& 38.842 & 16.781 & 8.434 & 4.180 & 1.020\\
skew-$t$	& 1 & 0	& 31.845 & 14.542 & 7.533 & 3.645 & {\bf 0.844}\\
skew-$t$	& 1 & 0.9	& 32.132 & 14.296 & 7.484 & 3.497 & 0.890\\
skew-$t$	& 3 & 0	& 33.653 & 15.453 & 8.119 & 4.338 & 1.188\\
skew-$t$	& 3 & 0.9 	& 32.157 & 14.727 & 7.794 & 3.825 & 0.917\\
\hline
    \end{tabular}
    \caption{Brier score for predicting exceedance of $c = \hat{q}(\tau)$ from five-fold cross-validation ($\times 1000$)}
  \end{table} \vspace{-1.5em}
  \begin{itemize}
  	\item Quantile score results are similar
  \end{itemize}
\end{frame}

% % TODO: Update with new plots
% \begin{frame}{Predicted 95th quantile}
% \centering
% \begin{figure}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-gau.pdf}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
%     \caption{Predicted 95th quantile using Gaussian and $t$}
% \end{figure}
% \end{frame}

% \begin{frame}{Predicted 95th quantile}
% \centering
% \begin{figure}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-t19.pdf}
%     \caption{Predicted 95th quantile using $t$ and $t$ thresholded at $T=0.9$}
% \end{figure}
% \end{frame}

% \begin{frame}{Predicted 99th quantile}
% \centering
% \begin{figure}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-gau.pdf}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
%     \caption{Predicted 99th quantile using Gaussian and $t$}
% \end{figure}
% \end{frame}

% \begin{frame}{Predicted 99th quantile}
% \centering
% \begin{figure}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-t19.pdf}
%     \caption{Predicted 99th quantile using $t$ and $t$ thresholded at $T=0.9$}
% \end{figure}
% \end{frame}

% \begin{frame}{Probability of exceedance}
% \centering
% \begin{figure}
%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-gau.pdf}
%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
%     \caption{Probability of exceeding the 75 ppb ozone standard using Gaussian and $t$}
% \end{figure}
% \end{frame}

% \begin{frame}{Probability of exceedance}
% \centering
% \begin{figure}
%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t19.pdf}
%     \caption{Probability of exceeding the 75 ppb ozone standard using $t$ and $t$ thresholded at $T=0.9$}
% \end{figure}
% \end{frame}
